Supplementary Information

### 2020-09-01 doc/manuscript/MEE_manuscript.Rmd
added exclusion of records with obvious errors

All tree cores (n=*#*) were measured and cross-dated by the original researchers using standard dendrochronological practices. 
The full record for all cores was retained for analysis, with the exception of records that had to be excluded due to insufficient sample size or anomalous growth patterns.
Specifically, we excluded
(1) species with <7 cores,
(2) cores with <30 years of record,
(3) contiguous portions of cores containing large outliers ($\Delta r$ > mean plus 5 x SD of $\Delta r$ for the entire core),
and (4) the final 20 years prior to death of trees cored dead. 
The final criteria was implemented to avoid periods of growth decline and potentially altered climate sensitivity prior to death [@cailleret_synthesis_2017; @desoto_low_2020].
From analyses including $DBH$ (see below), we further excluded
(1) trees for which we lacked data required to reconstruct $DBH$,
(2) trees for which there was a significant inconsistency between measured $DBH$ and the sum of $\Delta r$'s across the core (*Appendix S1*),
and (3) parts of records where reconstructed $DBH$ (see below) was represented by <3 conspecific trees.
In total, this resulted in inclusion of *#* cores, *#* of which could be included in analyses with $DBH$ (Table S3).

>>>>

All tree cores were measured and cross-dated by the original researchers using standard dendrochronological practices. 
From among the full set of *#* original records, we excluded cores for which we detected errors (e.g., labeling inconsistencies, obvious dating errors) that could not be resolved before finalizing the analysis (n=*#*). 
We also excluded records that had to be excluded due to insufficient sample size or anomalous growth patterns, including
(1) species with <7 cores,
(2) cores with <30 years of record,
(3) contiguous portions of cores containing large outliers ($\Delta r$ > mean plus 5 x SD of $\Delta r$ for the entire core),
and (4) the final 20 years prior to death of trees cored dead. 
The final criteria was implemented to avoid periods of growth decline and potentially altered climate sensitivity prior to death [@cailleret_synthesis_2017; @desoto_low_2020].
From analyses including $DBH$ (see below), we further excluded
(1) trees for which we lacked data required to reconstruct $DBH$,
(2) trees for which there was a significant inconsistency between measured $DBH$ and the sum of $\Delta r$'s across the core (*Appendix S1*),
and (3) parts of records where reconstructed $DBH$ (see below) was represented by <3 conspecific trees.
In total, this resulted in inclusion of *#* cores, *#* of which could be included in analyses with $DBH$ (Table S3).
